Enabling GPU Virtualization in Cloud Environments

### 2015-10-23 paper.tex
motivation revisado

Cloud computing, storage and other facilities are becoming 
more popular during the last years thanks to the easy set up 
process and the low required IT invest. The features provided 
by the cloud services are useful for companies and individuals 
which have access to the last technology without spending a 
huge amount of money. Moreover, the systems offered by the cloud 
companies matches with the ones required thanks to its flexibility.

Two different kinds of cloud approaches are available:
on the one hand, public facilities such as Amazon Web Services or 
Microsoft Azure that allow users to chose a pre-configured system  
and on the other hand, institutions/companies 
are able to deploy their own cloud infrastructure by using software products 
like OpenStack.

In both cases, the power of current cloud facilities is based on the use 
of Virtual Machines (VM) which  
emulates a physical node with its own 
CPUs, storage, network interconnection, OS and GPU. 
However, to provide a VM with GPUs means
that the guest host must be equipped with at least one GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VM that are allowed to use GPUs is 
limited by the number of the installed devices. This lack of 
flexibility is also noticed in the cluster configurations, where 
depending on the users’ necessities, 
two approaches to introduce GPGPU computation can be determined:
the conservative, to have a lot of hosts with GPUs to be
ensured that you are likely to satisfy all the requests; or a
more daring approach, where you will count with some GPU-
enabled nodes hoping not to arrive a large burst of requests
so the cluster would run out of GPU resources.

The use of a virtualization technology such as {rCUDA} adds 
the desired flexibility because this software detaches the GPUs 
from the nodes where they are installed allowing all nodes in a cluster
not only use them as if they were local but also to share them 
achieving a higher utilization rate and reducing the idle status time.
Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
environment should be able to use and share all the GPUs installed in 
the physical nodes removing the current limitations.

Our work have been focused on Amazon Web Service as example of public 
solution and on OpenStack software acting as a private cloud environment. 
In order to achieve it, these cloud facilities need to be provided with the {rCUDA} software and 
in the case of OpenStack, a module for managing virtualized GPUs 
requires to be implemented.

>>>>

Cloud computing and storage have gained wide appeal  
during the last years thanks to the easy set up 
process and the low IT invest. The services offered in the cloud are useful for companies and individuals 
that look for temporary access to the last technology without spending a 
huge amount of money. In general, the systems provided by cloud 
companies match with the those requested thanks to its flexibility.

Two different kinds of cloud approaches are currently available:
on one hand, public facilities, such as AWS or 
Microsoft Azure that allow users to chose a pre-configured system.   
On the other hand, institutions/companies 
are able to deploy their own cloud infrastructure by a software as OpenStack.

In both cases, the power of current cloud facilities relies on the use 
of VMs which  
emulate a physical node with its own 
CPUs, storage, network interconnection, GPU and operating system (OS). 
However, providing a VM equipped with a GPU implies
that the guest host must be equipped with at least one GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VMs that are allowed to use GPUs is 
limited by the number of the installed devices. This lack of 
flexibility is also present in the cluster configurations, where 
depending on the users' necessities, 
two approaches to introduce GPGPU computation can be identified:
the conservative option, features a lot of hosts with GPUs to 
ensure that is likely to satisfy all requests; in a
more daring approach, a reduced count of some GPU-enabled nodes is available, in the hope that a large burst of GPU requests 
 does not exhaust the cluster's resources.

The use of a virtualization technology such as {rCUDA} adds 
the desired flexibility because this software detaches the GPUs 
from the nodes where they are installed, allowing all nodes in a cluster
not only to use them as if they were local, but also to share them 
achieving a higher utilization rate and reducing the idle periods.
Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
environment can leverage and share all GPUs installed in 
the physical nodes, removing past limitations.

Our work focuses on AWS as a public 
solution, and on OpenStack software as a private cloud environment. 
In order to achieve our goal, these cloud facilities need to be provided with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
needs to be implemented.
